---
title: DATA 606 Data Project Proposal
author: "Ohannes (Hovig) Ohannessian"
date: "4/8/2018"
output: html_document
---

### Data Preparation

```{r echo=FALSE, include=FALSE}
library(tidyr)
library(dplyr)
library(knitr)
library(rworldmap)
library(ggplot2)
library(ggmap)
library(plotly)
```

<div class="alert alert-success">
<b><u>City of Chicago - Locations of Array of Things sensor nodes</u></b><br>
Listing an overview of projects and plans that some went live and some are still planned and locating sensors on the map.
</div>

```{r}
array_of_things_locations_data<-read.csv("https://raw.githubusercontent.com/hovig/MSDS_CUNY/master/DATA606/Project-Proposal/array-of-things-locations-1.csv")

status_of_things<-array_of_things_locations_data %>% 
                    group_by(Status) %>%
                    summarise(count=n())

dat <- data.frame(
    status = factor(status_of_things$Status, levels=status_of_things$Status),
    count = status_of_things$count
)

df<-round(data.frame(
  x = jitter(array_of_things_locations_data$Longitude, amount = .3),
  y = jitter(array_of_things_locations_data$Latitude, amount = .3)), 
  digits = 2)
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Whats the status of plans for the Array of Things project done by the City of Chicago? And where the sensors are located?


### Cases 

**What are the cases, and how many are there?**

Each case represents a sensor. There are `r nrow(array_of_things_locations_data)` observations in the given data set.


### Data collection 

**Describe the method of data collection.**

- For more information on the Array of Things project, see https://arrayofthings.github.io.
- Last updated at https://data.cityofchicago.org/ : 2018-02-27


### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

<u>The data is collected by the City of Chicago and found here for more relevance:</u>

- Source: https://data.cityofchicago.org/d/6rq2-yx28


### Response 

**What is the response variable, and what type is it (numerical/categorical)?**

The response variable is status and is categrical.


### Explanatory 

**What is the explanatory variable, and what type is it (numerical/categorival)?**

The explanatory variables are the count and the geolocations and are numerical.


### Relevant summary statistics 

**Provide summary statistics relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**


```{r}
glimpse(array_of_things_locations_data)
```

```{r}
kable(status_of_things)
```

```{r}
ggplot(data=dat, aes(x=status, y=count, fill=time)) + 
    geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") + 
    guides(fill=FALSE) +
    xlab("Status type") + ylab("Status count per type") +
    ggtitle("Chicago's planning status")
```

```{r echo=FALSE, include=FALSE}
map<-get_googlemap('chicago', markers = df, path = df, scale = 2)
```

```{r}
ggmap(map, extent = 'device')
```
