---
title: "HW2 - SQL + R - Data607"
author: "Ohannes (Hovig) Ohannessian"
date: "2/8/2018"
output: 
  html_document: default
  pdf_document: default
---

```{r echo = FALSE, message = FALSE}
library(sqldf)
library(RMySQL)
library(DBI)
library(ggplot2)
```




#### **_<u>Assignment Question</u>_:**


**Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.**


# Movie Ratings

```{r echo = FALSE}
db <- dbConnect(MySQL(), user='root', host='localhost', dbname='movies')
print(paste("The 2 tables in movies MySQL database (",dbListTables(db)[1],"and",dbListTables(db)[2],") are inner joined and sorted in descending order by the rating value column."))
```

## Chosen Movie List 

```{r, echo = FALSE}
library(knitr)
```


### <u>Table</u>

```{r echo = FALSE, results = 'asis'}
query<-"select m.movie_id, movie_title, rating_value, rating_owner from movies as m inner join ratings where m.movie_id = ratings.movie_id order by rating_value desc;"
kable(fetch(dbSendQuery(db,query)))
```

### <u>Analysis</u>

##### **_Summarise the db table_**
```{r echo = FALSE, results = 'asis'}
df <- fetch(dbSendQuery(db,query))
kable(summary(df))
```

##### **_Split duplicates with their ratings_**
```{r echo = FALSE}
val<-c(split(df$rating_value, duplicated(df$movie_title) | duplicated(df$movie_title, fromLast = FALSE)),split(df$movie_title, duplicated(df$movie_title) | duplicated(df$movie_title, fromLast = FALSE)))
val1<-as.double(unlist(val[1][1]))
val2<-as.double(unlist(val[2][1]))
val3<-unlist(val[3][1])
val4<-unlist(val[4][1])
dfval<-data.frame(val1,val3,val2,val4)
Rating_1of2<-dfval$val1
Movies_1of2<-dfval$val3
Rating_2of2<-dfval$val2
Movies_2of2<-dfval$val4
dfval<-data.frame(Rating_1of2,Movies_1of2,Rating_2of2,Movies_2of2)
kable(dfval)
```

##### **_Add ratings of duplicates to list high and low rated movies_**
```{r echo = FALSE}
vec<-c()
for(i in 1:length(val1)) {
  vec[i]<-sum(val1[i]+val2[i])
}

score<-c()
for(j in 1:length(vec)){
  if(vec[j]==max(vec)){
    Highest_Score<-vec[j]
  }
  if(vec[j]==min(vec)){
    Lowest_Score<-vec[j]
  }
}

dfscore<-data.frame(movies=c(highest_score=val3[match(c(Highest_Score),vec)],lowest_score=val3[match(c(Lowest_Score),vec)]))
Highest_rated_movie<-dfscore$movies[1]
Lowest_rated_movie<-dfscore$movies[2]
kable(data.frame(Highest_rated_movie,Lowest_rated_movie))
```

### <u>Plot the db table</u>

```{r echo = FALSE, results = 'asis'}
Movies<-df$movie_title
Ratings<-df$rating_value
Raters<-df$rating_owner
ggplot(df, aes(x=Movies, y=Ratings)) + geom_point(shape=1) + geom_smooth(method=lm , color="red", se=TRUE)
```

