---
title: "Lab 8 - Data 606"
author: "Ohannes (Hovig) Ohannessian"
date: "5/1/2018"
output: html_document
---
<hr>
## Grading the professor
```{r include=FALSE}
library(vcd)
library(ggplot2)
```

### The data

```{r load-data, eval=TRUE}
load(url("http://www.openintro.org/stat/data/evals.RData"))
```


### Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.

<div class="alert alert-info">
* It is an observational study. 
* How non-teaching related characteristics (mainly the physical appearance) influence on student's evaluation of the course?
</div>

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?

```{r}
hist(evals$score)
summary(evals$score)
```

<div class="alert alert-info">
* the distribution appears left skewed
* the mean is `r summary(evals$score)[4]`. The rates are more positive.
* More information is needed to predict what to expect, but students' ratings seem to be high and satisfactory for them.
</div>

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).

```{r}
mosaic(~ evals$ethnicity + evals$gender)
```

<div class="alert alert-info">
* minority: the # female professors is greater than the male
* non-minority: the # female professors is less than the male 
</div>

### Simple linear regression


```{r}
plot(evals$score ~ evals$bty_avg)
```


4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?
    
```{r}
qplot(x = evals$bty_avg, y = evals$score, geom = "point") + geom_jitter()
```

<div class="alert alert-info">
* the points were more linear in the initial scatterplot
* with `jitter`, the scatterplot displays points in a cloud-looking cluster of points
</div>

5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?

```{r}
m_bty<-lm(evals$score ~ evals$bty_avg)
plot(jitter(evals$score) ~ jitter(evals$bty_avg))
abline(m_bty)

summary(m_bty)
```

<div class="alert alert-info">
* $\hat{y} = 3.88034 + 0.06664 \times$ bty_avg
* somewhat predictive with high residuals standard error of 53.48% and $R^2_{adj}$ = 3.293%
</div>

6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).

```{r}
plot(m_bty$residuals ~ evals$bty_avg)
abline(h = 0)

hist(m_bty$residuals)

qqnorm(m_bty$residuals)
qqline(m_bty$residuals)
```

<div class="alert alert-info">
* Linearity: the residual plot is randomly dispersed, but satisfied, linearity is met
* Normal residuals: left skewed histogram rather than normal distribution. The almost "S" shape indicates the non-normality of the residuals
* Constant variability: it is fairly constant
* Independence of observations: we assume the randomness and the indepence of the observations.
</div>

### Multiple linear regression



```{r}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```



```{r}
plot(evals[,13:19])
```



```{r}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.

```{r}
plot(m_bty_gen$residuals ~ evals$bty_avg)
abline(h = 0)

hist(m_bty_gen$residuals)

qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals)
```

<div class="alert alert-info">
* Linearity: the residual plot is randomly dispersed, but satisfied, linearity is met
* Normal residuals: left skewed histogram rather than normal distribution. The almost "S" shape indicates the non-normality of the residuals
* Constant variability: it is fairly constant
* Independence of observations: we assume the randomness and the indepence of the observations.
</div>

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?

```{r}
summary(m_bty) 
summary(m_bty_gen)
```

<div class="alert alert-info">
* bty_avg is still the significant predictor 
* gender addition didn't effect much, it only helped $R^2$ and $R^2_{adj}$ to slightly improve
* Residual standard error = 0.5287 didn't change much 
</div>


As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?

<div class="alert alert-info">
* gendermale = 1
* $\hat{y} = 3.74734 + 0.07416 \times$ bty_avg + 0.17239 = 3.91973 + 0.07416 $\times$ bty_avg 
* $\hat{y}_{male}$ > $\hat{y}_{female}$
</div>



10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.

```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data=evals)
summary(m_bty_rank)
```

<div class="alert alert-info">
* For n levels, R considers n-1 variables or n-1 categories. 
</div>


11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.

```{r}
variables<-c("rank" , "ethnicity" , "gender" , "language" , "age" , "cls_perc_eval" , "cls_students" , "cls_level" , "cls_profs" , "cls_credits" , "bty_avg" , "pic_outfit" , "pic_color")

cor_results<-c(cor(as.numeric(evals$score),as.numeric(evals$rank)),cor(as.numeric(evals$score),as.numeric(evals$ethnicity)),cor(as.numeric(evals$score),as.numeric(evals$gender)),cor(as.numeric(evals$score),as.numeric(evals$language)),cor(as.numeric(evals$score),as.numeric(evals$age)),cor(as.numeric(evals$score),as.numeric(evals$cls_perc_eval)),cor(as.numeric(evals$score),as.numeric(evals$cls_students)),cor(as.numeric(evals$score),as.numeric(evals$cls_level)),cor(as.numeric(evals$score),as.numeric(evals$cls_profs)),cor(as.numeric(evals$score),as.numeric(evals$cls_credits)),cor(as.numeric(evals$score),as.numeric(evals$bty_avg)),cor(as.numeric(evals$score),as.numeric(evals$pic_outfit)),cor(as.numeric(evals$score),as.numeric(evals$pic_color)))

for(i in cor_results){
  if(i>=-0.026 & i<=0.026){
      print(variables[which(cor_results == i)])
  }
}
```

<div class="alert alert-info">
* cls_profs is the closest to have no relationship (almost 0)
</div>


```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.

<div class="alert alert-info">
* cls_profssingle has the highest p-value of 0.77806 as we confirmed it previously by choosing cls_profs
</div>

13. Interpret the coefficient associated with the ethnicity variable.

<div class="alert alert-info">
* `ethnicitynot minority` variable has a coefficent of 0.1234929 that is a multiplier to itself 
</div>

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?

```{r}
without_high_p_value <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + cls_students + cls_level + cls_credits + bty_avg + pic_outfit + pic_color, data = evals)
summary(without_high_p_value)
```

<div class="alert alert-info">
* we notice minor changes in coefficients of the variables
* we notice a decrease in the p-value
* we notice an increase in the $R^2_{adj}$
</div>

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.

```{r}
backward_selection <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval + 
    cls_credits + bty_avg + pic_color, data = evals)
summary(backward_selection)
```

<div class="alert alert-info">
* $\hat{y}$ = 3.771922 + 0.167872 $\times$ ethnicity_not_minority + 0.207112 $\times$ gender_male - 0.206178 $\times$ language_non_english 
              - 0.006046 $\times$ age + 0.004656 $\times$ cls_perc_eval + 0.505306 $\times$ cls_credits_one_credit + 0.051069 $\times$ bty_avg 
              - 0.190579 $\times$ pic_color_color
</div>

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.

```{r}
hist(backward_selection$residuals)

qqnorm(backward_selection$residuals)
qqline(backward_selection$residuals)

plot(backward_selection$residuals)
abline(h = 0)
```

<div class="alert alert-info">
* Linearity: the residual plot is randomly dispersed, but satisfied, linearity is met
* Normal residuals: left skewed histogram rather than normal distribution. The almost "S" shape indicates the non-normality of the residuals
* Constant variability: it is fairly constant
* Independence of observations: we assume the randomness and the indepence of the observations.
</div>

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?

<div class="alert alert-info">
* adding a course will impact on the conditions of linear regression
</div>

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.

<div class="alert alert-info">
Characteristics would be:

* a non-minority male<br>
* who speaks english<br>
* is younger<br>
* has a high percentage of students complete their evaluation<br>
* is considered attractive<br>
* has a more casual picture in color
</div>

19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?


<div class="alert alert-info">
Better not generalize it because:

* this is one observational study in one university<br>
* the definition of beauty may not be the case elsewhere
</div>
