---
title: "Lab4a - Data 606"
author: "Ohannes (Hovig) Ohannessian"
date: "3/17/2018"
output: html_document
---


```{r load-data, eval=TRUE}
library(knitr)
load(url("http://www.openintro.org/stat/data/ames.RData"))
```


```{r assign, eval=TRUE}
area <- ames$Gr.Liv.Area
price <- ames$SalePrice
```


```{r area, eval=TRUE}
summary(area)
hist(area)
```

1.  Describe this population distribution.
<div class="alert alert-warning">
The population distribution is skewed to the right and the shape of the distribution is in between the `r min(area)` and `r max(area)`.
</div>

```{r samp1, eval=TRUE}
samp1 <- sample(area, 50)
```


2.  Describe the distribution of this sample. How does it compare to the distribution of the population?
<div class="alert alert-warning">
This sample also is right skewed.
</div>


```{r mean-samp1, eval=TRUE}
mean(samp1)
```



3.  Take a second sample, also of size 50, and call it `samp2`.  How does the 
    mean of `samp2` compare with the mean of `samp1`?  Suppose we took two 
    more samples, one of size 100 and one of size 1000. Which would you think 
    would provide a more accurate estimate of the population mean?
```{r}
mu1<-mean(samp1)
samp2<-sample(area,50)
mu2<-mean(samp2)
if(mu1<mu2){
  printo<-paste("The mean of samp2 is greater than samp1")
} else if(mu1>mu2){
  printo<-paste("The mean of samp2 is less than samp1")
} else {
  printo<-paste("The mean of samp2 is equal to samp1")
}
```

<div class="alert alert-warning">
The mean of samp2 is `r mean(samp2)` vs. the mean of samp1 is `r mean(samp1)`. `r printo`.
</div>


```{r loop, eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   }

hist(sample_means50)
```


```{r hist-breaks, eval=TRUE}
hist(sample_means50, breaks = 25)
```


4.  How many elements are there in `sample_means50`?  Describe the sampling 
    distribution, and be sure to specifically note its center.  Would you 
    expect the distribution to change if we instead collected 50,000 sample 
    means?
<div class="alert alert-warning">
\# of elements of `sample_means50`: `r length(sample_means50)` with a mean of: `r round(mean(sample_means50),2)`. The distribution looks normal.
</div>

```{r loop-long, eval=TRUE}
sample_means50 <- rep(NA, 5000)

samp <- sample(area, 50)
sample_means50[1] <- mean(samp)

samp <- sample(area, 50)
sample_means50[2] <- mean(samp)

samp <- sample(area, 50)
sample_means50[3] <- mean(samp)

samp <- sample(area, 50)
sample_means50[4] <- mean(samp)
```


```{r eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   if(i==1){
    print("from ")
    print(i)
    print(" to ")
   }
   if(i==5000)
    print(i)
   }
```


5.  To make sure you understand what you've done in this loop, try running a 
    smaller version.  Initialize a vector of 100 zeros called 
    `sample_means_small`.  Run a loop that takes a sample of size 50 from 
    `area` and stores the sample mean in `sample_means_small`, but only 
    iterate from 1 to 100.  Print the output to your screen (type 
    `sample_means_small` into the console and press enter).  How many elements 
    are there in this object called `sample_means_small`? What does each 
    element represent?

```{r loop-again, eval=TRUE}
sample_means_small <- rep(NA, 100)

for(i in 1:100){
   samp <- sample(area, 50)
   sample_means_small[i] <- mean(samp)
   if(i==1){
    print("from ")
    print(i)
    print(" to ")
   }
   if(i==100)
    print(i)
}
```

<div class="alert alert-warning">
\# of elements of `sample_means_small`: `r length(sample_means_small)` with a mean of: `r round(mean(sample_means_small),2)`. Each element represents the mean of the sample of 50, rolled 100 times.
</div>

```{r hist, eval=TRUE}
hist(sample_means50)
```


```{r samp-10-100, eval=TRUE}
sample_means10 <- rep(NA, 5000)
sample_means100 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(area, 10)
  sample_means10[i] <- mean(samp)
  samp <- sample(area, 100)
  sample_means100[i] <- mean(samp)
}
```



```{r plot-samps, eval=TRUE, tidy = FALSE}
par(mfrow = c(3, 1))

xlimits <- range(sample_means10)

hist(sample_means10, breaks = 20, xlim = xlimits)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means100, breaks = 20, xlim = xlimits)
```


6.  When the sample size is larger, what happens to the center?  What about the spread?

<div class="alert alert-warning">
* larger the sample size is, the closer to the main population is
* the sample mean gets closer to the population mean
* the spread becomes shorter
</div>

* * *
## On your own

So far, we have only focused on estimating the mean living area in homes in 
Ames.  Now you'll try to estimate the mean home price.

-   Take a random sample of size 50 from `price`. Using this sample, what is 
    your best point estimate of the population mean?

```{r}
sample_price50<-sample(price,50)
mean_sample_price50<-mean(sample_price50)
point_estimate_sample_price50<-format(sample_price50,scientific=FALSE)
```

<div class="alert alert-warning">
* sample_price_of_50 = `r sample_price50`
* with a mean of: mean_sample_price_of_50 = `r round(mean_sample_price50,2)` 
* with a point estimate of: point_estimate_sample_price_of_50 = `r point_estimate_sample_price50`.
</div>

-   Since you have access to the population, simulate the sampling 
    distribution for $\bar{x}_{price}$ by taking 5000 samples from the 
    population of size 50 and computing 5000 sample means.  Store these means 
    in a vector called `sample_means50`. Plot the data, then describe the 
    shape of this sampling distribution. Based on this sampling distribution, 
    what would you guess the mean home price of the population to be? Finally, 
    calculate and report the population mean.

```{r}
sample_means50<-rep(NA, 5000)
for(i in 1:5000){
  samp<-sample(price, 50)
  sample_means50[i]<-mean(samp)
}
hist(sample_means50, breaks=25)
```

<div class="alert alert-warning">
* The mean of `sample_means50` is between 180000 and 181000. mean_sample_means50 = `r mean(sample_means50)`
* The shape of `sample_means50` distribution is normal and at the `mean_sample_means50`
* The point estimate is: point_estimate_sample_means50 = `r format(mean(sample_means50), scientific=FALSE)`
</div>

-   Change your sample size from 50 to 150, then compute the sampling 
    distribution using the same method as above, and store these means in a 
    new vector called `sample_means150`. Describe the shape of this sampling 
    distribution, and compare it to the sampling distribution for a sample 
    size of 50.  Based on this sampling distribution, what would you guess to 
    be the mean sale price of homes in Ames?

```{r}
sample_means150<-rep(NA, 5000)
for(i in 1:5000){
  samp<-sample(price, 150)
  sample_means150[i]<-mean(samp)
}
```

<div class="alert alert-warning">
* The mean of `sample_means150` is between 180000 and 181000. mean_sample_means150 = `r mean(sample_means150)`
* The shape of `sample_means150` distribution is normal and at the `mean_sample_means150`
* The point estimate is: point_estimate_sample_means150 = `r format(mean(sample_means150), scientific=FALSE)`
* The comparison will show that larger the sample is, better will be the shape of the distribution
</div>

```{r}
par(mfrow = c(1, 2))
xlimits <- range(sample_means50)

hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means150, breaks = 20, xlim = xlimits)
```

-   Of the sampling distributions from 2 and 3, which has a smaller spread?  If
    we're concerned with making estimates that are more often close to the 
    true value, would we prefer a distribution with a large or small spread?
```{r}
xlimits50 <- range(sample_means50)
range50diff <- xlimits50[2] - xlimits50[1]

xlimits150 <- range(sample_means150)
range150diff <- xlimits150[2] - xlimits150[1]
```

<div class="alert alert-warning">
* The range of `sample_means50` is `r xlimits50`
* The range of `sample_means150` is `r xlimits150`
* range50diff = `r range50diff`
* range150diff = `r range150diff`
* the spread becomes shorter when the sample becomes larger: `range150diff` < `range50diff`
</div>



