---
title: "Final Project - CUNY Data 607"
author: "Ohannes (Hovig) Ohannessian"
date: "4/19/2018"
output: html_document
---
<hr>
## <u>Hockey Teams' Points Based Ranking Predictions</u>

Check project's: 

* [Proposal](http://rpubs.com/hovig613/383339)
* [Presentation](http://rpubs.com/hovig613)

 	
```{r setup, include=FALSE}
library(ggplot2) 
library(knitr)
library(psych) 
```

<br><div class="alert alert-info">
* Load train & test datasets from final project Github repo into variables
</div>

```{r}
train_url<-"https://raw.githubusercontent.com/hovig/MSDS_CUNY/master/Data607/final_project/datasets/final_project_train.csv"
train<-read.csv(train_url)

test_url<-"https://raw.githubusercontent.com/hovig/MSDS_CUNY/master/Data607/final_project/datasets/final_project_test.csv"
test<-read.csv(test_url)
```

### <u>Train</u>

<div class="alert alert-info">
* Train dataset: display content information
</div>

```{r}
str(train)
```

<br><div class="alert alert-info">
* Train dataset: column titles
</div>

```{r}
kable(data.frame(column_names=names(train)))
```

<br><div class="alert alert-info">
* Train dataset: scale, correlate and plot
</div>

```{r}
train_scale<-scale(train[, -1:-2])
correlation<-cor(train_scale)
cor.plot(correlation)
```

<br><div class="alert alert-info">
* Apply principal component analysis on train scaled dataset without rotation 
* Plot this analysis
</div>

```{r}
pca_rotate_not<-principal(train_scale, rotate="none")
plot(pca_rotate_not$values, type="b", ylab="Eigenvalues", xlab="Component")
```

<br><div class="alert alert-info">
* Apply Principal component analysis on train scaled dataset with rotation 
* Put it in data frame
* Display parts of the score data frame
</div>

```{r}
pca_rotate<-principal(train_scale, nfactors = 5, rotate = "varimax")
pca_rotate

pca_scores<-data.frame(pca_rotate$scores)
head(pca_scores)
```

<br><div class="alert alert-info">
* Insert points per game values into scores data frame 
* Apply linear modeling on it `lm()`
* Display summaries of these linear model metrics
* Plot the predicted values with the original ones
</div>

```{r}
pca_scores$ppg<-train$ppg

linear_model_all<-lm(ppg~., data = pca_scores)
summary(linear_model_all)

linear_model_rc1_rc2<-lm(ppg~RC1+RC2, data = pca_scores)
summary(linear_model_rc1_rc2)
```

<br><div class="alert alert-info">
* scatterplot
</div>

```{r}
plot(linear_model_rc1_rc2$fitted.values, train$ppg, main="Predicted vs. Original", xlab="Predicted",ylab="Original")

train$predicted<-round(linear_model_rc1_rc2$fitted.values, digits = 2)

ggplot(train, aes(x = predicted, y = ppg, label = Team)) + 
  geom_point() + 
  geom_text(size=3.5, hjust=0.1, vjust=-0.5, angle=0) + 
  xlim(0.8, 1.4) + ylim(0.8, 1.5) +
  stat_smooth(method="lm", se=FALSE)

pca_scores$Team<-train$Team

ggplot(pca_scores, aes(x = RC1, y = RC2, label = Team)) + 
  geom_point() +
  geom_text(size = 2.75, hjust = .2, vjust = -0.75, angle = 0) +
  xlim(-2.5, 2.5) + ylim(-3.0, 2.5)
```

<br><div class="alert alert-info">
* Root Means Squared Error (RMSE): calculate error by finding the mean of residuals (from linear model) and square root that value
</div>

```{r}
sqrt(mean(linear_model_rc1_rc2$residuals^2))
```

<div class="alert alert-info">
* train_rmse = `r sqrt(mean(linear_model_rc1_rc2$residuals^2))` 
</div>
<hr>

### <u>Test</u>

<div class="alert alert-info">
* Apply `predict()` function to scale the test data
* Insert and predict the rotated and test data values into data frame 
* Insert to Test predicted scores the predicted values of linear model of Train, rotated and Test datasets
* Plot the Test predicted values with the points per game
</div>

```{r}
test_scores<-data.frame(predict(pca_rotate, test[, c(-1:-2)]))
test_scores$predicted<-predict(linear_model_rc1_rc2, test_scores)

test_scores$ppg<-test$ppg
test_scores$Team<-test$Team

ggplot(test_scores, aes(x = predicted, y = ppg, label = Team)) + 
  geom_point() + 
  geom_text(size=3.5, hjust=0.4, vjust = -0.9, angle = 35) + 
  xlim(0.75, 1.5) + ylim(0.5, 1.6) +
  stat_smooth(method="lm", se=FALSE)
```

<br><div class="alert alert-info">
* Root Means Squared Error (RMSE): find the mean of residuals by subtracting predicted values from ppg in Test scores data frame and square root that value
</div>

```{r}
residuals<-test_scores$ppg - test_scores$predicted
sqrt(mean(residuals^2))
```

<div class="alert alert-info">
* test_rmse = `r sqrt(mean(residuals^2))` 
</div>

### <u>Summary</u>
<div class="alert alert-info">
> In this project, 

> * we examined and applied the Process component analysis to unsupervised learning. 
> * We reduced the dimensions to understand the dataset in a supervised learning and 
> * we used the linear regression analysis to make prediction of each team's total points per game.
> * We used visualization to plot the data and the pricipal components.
> * With more training needed to better the prediction and reduce error but `test_rmse > train_rmse` validates our model.
```markdown
Our points per game modeling shows the team rankings of:
  - Colorado Avalanche as the worst  
  - Washington Capitals as the best
```
</div>

### <u>Useful Links</u>

* <a href="http://www.nhl.com/stats" " target="_blank">NHL Statistics</a>
* <a href="https://www.hockey-reference.com/" " target="_blank">Hockey Reference</a>
* <a href="http://www.opensourcesports.com/hockey/" " target="_blank">Open Source Sports - Hockey</a>
* <a href="https://en.wikipedia.org/wiki/Principal_component_analysis" " target="_blank">Principal component analysis</a>
* <a href="https://raw.githubusercontent.com/hovig/MSDS_CUNY/master/Data607/final_project/" " target="_blank">Final Project on Github</a><br>





