---
title: "Lab 5 - Data 606"
author: "Ohannes (Hovig) Ohannessian"
date: "3/25/2018"
output: html_document
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis


```{r load-data, eval=TRUE}
library(knitr)
load(url("http://www.openintro.org/stat/data/nc.RData"))
```


1.  What are the cases in this data set? How many cases are there in our sample?

```{r summary, eval=TRUE}
summary(nc)
```

<div class="alert alert-success">
The cases in this data set are the birth observations in North Carolina. There are `r nrow(nc)` cases.
</div>

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r}
boxplot(weight~habit,data=nc, main="Mother's Habit vs. Baby's Weight", ylab="Weight", xlab="Habit")
```
<div class="alert alert-success">
The median birth weight of babies born to non-smoker mothers is slightly higher than that of babies born to smoker mothers.
</div>

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r eval=TRUE}
by(nc$weight, nc$habit, length)
```

<div class="alert alert-success">
* <u>Conditions:</u><br>
1- Random: assuming randomness is in place<br>
2- Normal: sample sizes are large enough to be normal with central limit sampling distribution >30<br>
3- Independent: births do not depend on one another 
</div>

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

<div class="alert alert-success">
$H_0: \mu_{smoking} = \mu_{non-smoking}$<br>
$H_A: \mu_{smoking} \ne \mu_{non-smoking}$
</div>

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")
```

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

<div class="alert alert-success">
95% confidence interval is 0.0534 to 0.5777
</div>


```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, alternative = "twosided", method = "theoretical")
```


```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, alternative = "twosided", method = "theoretical", order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r}
inference(y = nc$weeks,est = "mean",conflevel = 95,type = "ci",alternative = "twosided",method = "theoretical")
```

<div class="alert alert-success">
95 % Confidence interval = ( 38.1528 , 38.5165 )
</div>

-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

```{r}
inference(y = nc$weeks,est = "mean",conflevel = 90,type = "ci",alternative = "twosided",method = "theoretical")
```

<div class="alert alert-success">
90 % Confidence interval = ( 38.182 , 38.4873 )
</div>

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

```{r}
inference(y=nc$weight, x=nc$mature, type="ht", est="mean", null=0, method="theoretical", alternative="twosided")
```

<div class="alert alert-success">
In the mean of the birth weight between younger and mature mothers:
There's no difference: $H_0: \mu_{younger} = \mu_{mature}$<br>
There's a difference: $H_A: \mu_{younger} \ne \mu_{mature}$

* p-value =  0.8526 > 0.05 => failed to reject $H_0$, therefore there's a similar birth weight between younger and mature mothers
</div>

-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

<div class="alert alert-success">
* the age cutoff for younger mothers <= `r max(nc[nc$mature == "younger mom",]$mage)`
* the age cutoff for mature mothers > `r min(nc[nc$mature == "mature mom",]$mage)`
</div>

-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

```{r}
inference(y=nc$visits, x=nc$marital, type="ht", est="mean", null=0, method="theoretical", alternative="twosided")
```

<div class="alert alert-success">
$H_0: \mu_{married_visits} = \mu_{non_married_visits}$<br>
$H_A: \mu_{married_visits} \ne \mu_{non_married_visits}$<br>
* p-value =  0 < 0.05 => fail to reject $H_A$, there's a difference in the average of doctor's visits between married and non-married mothers.
</div>
