---
title: "Lab 7 - Data 607"
author: "Ohannes (Hovig) Ohannessian"
date: "4/21/2018"
output: html_document
---

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
```

## Batter up 

The movie [Moneyball](http://en.wikipedia.org/wiki/Moneyball_(film)) focuses on
the "quest for the secret of success in baseball". It follows a low-budget team, 
the Oakland Athletics, who believed that underused statistics, such as a player's 
ability to get on base, betterpredict the ability to score runs than typical 
statistics like home runs, RBIs (runs batted in), and batting average. Obtaining 
players who excelled in these underused statistics turned out to be much more 
affordable for the team.

In this lab we'll be looking at data from all 30 Major League Baseball teams and
examining the linear relationship between runs scored in a season and a number 
of other player statistics. Our aim will be to summarize these relationships 
both graphically and numerically in order to find which variable, if any, helps 
us best predict a team's runs scored in a season.

## The data

Let's load up the data for the 2011 season.

```{r}
load(url("http://www.openintro.org/stat/data/mlb11.RData"))
```



1.  What type of plot would you use to display the relationship between `runs` 
    and one of the other numerical variables? Plot this relationship using the 
    variable `at_bats` as the predictor. Does the relationship look linear? If 
    you knew a team's `at_bats`, would you be comfortable using a linear model 
    to predict the number of runs?

<div class="alert alert-info">
* scatterplot
* linear relationship
* linear model can be used to predict the number of runs
</div>

```{r}
ggplot(mlb11, aes(x=at_bats, y=runs)) + geom_point()
```

If the relationship looks linear, we can quantify the strength of the
relationship with the correlation coefficient.

```{r cor, eval=TRUE}
cor(mlb11$runs, mlb11$at_bats)
```

## Sum of squared residuals

2.  Looking at your plot from the previous exercise, describe the relationship 
    between these two variables. Make sure to discuss the form, direction, and 
    strength of the relationship as well as any unusual observations.

<div class="alert alert-info">
* direction: positive
* form: linear, no strong curvature
* strength: weak, the points of the data are far spread from each other
</div>

```{r plotss-atbats-runs, eval=TRUE}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
```



```{r plotss-atbats-runs-squares, eval=TRUE}
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)
```



3.  Using `plot_ss`, choose a line that does a good job of minimizing the sum of
    squares. Run the function several times. What was the smallest sum of 
    squares that you got? How does it compare to your neighbors?

```{r}
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)
```

<div class="alert alert-info">
* Sum of Squares:  123721.9
* it's the same on every run  
* it's the same for all the neighbors
</div>


## The linear model



```{r m1, eval=TRUE}
m1 <- lm(runs ~ at_bats, data = mlb11)
```


```{r summary-m1, eval=TRUE}
summary(m1)
```


4.  Fit a new model that uses `homeruns` to predict `runs`. Using the estimates 
    from the R output, write the equation of the regression line. What does the 
    slope tell us in the context of the relationship between success of a team 
    and its home runs?

```{r}
plot_ss(x = mlb11$homeruns, y = mlb11$runs, showSquares = TRUE)
```

```{r}
cor(mlb11$runs, mlb11$homeruns)
summary(lm(runs ~ homeruns, data = mlb11))
```

<div class="alert alert-info">
${\beta}_0$ = 415.2389<br>
${\beta}_1$ = 1.8345<br>
y = ${\beta}_0$ + ${\beta}_1$ . x <br>=> runs = 415.2389 + 1.8345 * homeruns
</div>

## Prediction and prediction errors

Let's create a scatterplot with the least squares line laid on top.

```{r reg-with-line, eval=TRUE}
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1)
```


5.  If a team manager saw the least squares regression line and not the actual 
    data, how many runs would he or she predict for a team with 5,578 at-bats? 
    Is this an overestimate or an underestimate, and by how much? In other 
    words, what is the residual for this prediction?


```{r}
data.frame(mlb11$runs,mlb11$at_bats)
```

<div class="alert alert-info">
* given: at_bats = 5578
* predicted_runs = -2789.2429 + 0.6305 ∗ at_bats = `r (-2789.2429 + (0.6305 * 5578))`
* if we consider at_bats = 5579, its observed_runs = 713
* let's calculate the residual: e = observed_runs - predicted_runs = `r 713 - 727.6861`
* this means we overestimated the runs by `r abs(713 - 727.6861)`
</div>


## Model diagnostics


```{r residuals, eval=TRUE}
plot(m1$residuals ~ mlb11$at_bats)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

6.  Is there any apparent pattern in the residuals plot? What does this indicate
    about the linearity of the relationship between runs and at-bats?

<div class="alert alert-info">
* no pattern noticed in the residuals plot
* constant variability: satisfied because the residual plot shows how the spread is the same everywhere
</div>


*Nearly normal residuals*: To check this condition, we can look at a histogram

```{r hist-res, eval=TRUE}
hist(m1$residuals)
```

or a normal probability plot of the residuals.

```{r qq-res, eval=TRUE}
qqnorm(m1$residuals)
qqline(m1$residuals)  # adds diagonal line to the normal prob plot
```

7.  Based on the histogram and the normal probability plot, does the nearly 
    normal residuals condition appear to be met?

<div class="alert alert-info">
* the nearly normal residuals condition appear to be met
</div>

*Constant variability*:

8.  Based on the plot in (1), does the constant variability condition appear to 
    be met?
  
<div class="alert alert-info">
* constant variability: satisfied
</div>

* * *

## On Your Own

-   Choose another traditional variable from `mlb11` that you think might be a 
    good predictor of `runs`. Produce a scatterplot of the two variables and fit 
    a linear model. At a glance, does there seem to be a linear relationship?

```{r}
plot(x = mlb11$new_obs, y = mlb11$runs)
```

<div class="alert alert-info">
* strong linear relationship
</div>

-   How does this relationship compare to the relationship between `runs` and 
    `at_bats`? Use the R$^2$ values from the two model summaries to compare. 
    Does your variable seem to predict `runs` better than `at_bats`? How can you
    tell?
    
```{r}
summary(lm(runs ~ new_obs, data = mlb11))
```

```{r}
no<-summary(lm(runs ~ new_obs, data = mlb11))$r.squared
r<-summary(m1)$r.squared
```

<div class="alert alert-info">
* this relationship looks stronger compared to the relationship between `runs` and `at_bats`
* new_obs R^2 is: `r no` and runs R^2 is: `r r`
* new_obs is a better predictor for runs than to at_bats
</div>


-   Now that you can summarize the linear relationship between two variables, 
    investigate the relationships between `runs` and each of the other five 
    traditional variables. Which variable best predicts `runs`? Support your 
    conclusion using the graphical and numerical methods we've discussed (for 
    the sake of conciseness, only include output for the best variable, not all 
    five).

```{r}
plot(x = mlb11$bat_avg, y = mlb11$runs)
summary(lm(runs ~ bat_avg, data = mlb11))
```

<div class="alert alert-info">
By simply comparing each of the r-squared values, we find that "bat_avg", or batting average is the best predictor of runs scored, of the "traditional" variables.
</div>

-   Now examine the three newer variables. These are the statistics used by the 
    author of *Moneyball* to predict a teams success. In general, are they more 
    or less effective at predicting runs that the old variables? Explain using 
    appropriate graphical and numerical evidence. Of all ten variables we've 
    analyzed, which seems to be the best predictor of `runs`? Using the limited 
    (or not so limited) information you know about these baseball statistics, 
    does your result make sense?

```{r}
summary(lm(runs ~ new_onbase, data = mlb11))$r.squared
summary(lm(runs ~ new_slug, data = mlb11))$r.squared
summary(lm(runs ~ hits, data = mlb11))$r.squared
plot(mlb11$runs~mlb11$hits)
abline(lm(runs ~ hits, data = mlb11))
```

<div class="alert alert-info">
* these stats would be better predictors than the traditional stats
</div>

-   Check the model diagnostics for the regression model with the variable you 
    decided was the best predictor for runs.

```{r}
plot(m1$residuals ~ mlb11$new_obs)
abline(h = 0, lty = 3)

hist(m1$residuals)

qqnorm(m1$residuals)
qqline(m1$residuals) 
```

<div class="alert alert-info">
* No overwhelming pattern
* Histogram looks normal
* Normal probability plot of residuals looks good
</div>

