---
title: "Data 624 - Homework 1"
author: "Ohannes (Hovig) Ohannessian"
date: "2/10/2019"
output: html_document
---
<hr><hr>

# Chapter 2 - Time series graphics

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_packages}
library(fpp2)
library(forecast)
```

1. Use the help menu to explore what the series gold, woolyrnq and gas represent. These are available in the forecast package.

```{r warning=FALSE}
str(gold)
str(woolyrnq)
str(gas)
```

a) Use autoplot() to plot each of these in separate plots.

```{r warning=FALSE}
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
writeLines("")
```

b) What is the frequency of each series? Hint: apply the frequency() function.

```{r warning=FALSE}
cat("Gold Frequency: ", frequency(gold))
cat("Woolyrnq Frequency: ", frequency(woolyrnq))
cat("Gas Frequency: ", frequency(gas))
```

c) Use which.max() to spot the outlier in the gold series. Which observation was it?

```{r warning=FALSE}
cat("When gold got maximum value? ", which.max(gold))
cat("What was the gold's maximum value? ", gold[which.max(gold)])
```

<hr><hr>

2. Download the file tute1.csv from OTexts.org/fpp2/extrafiles/tute1.csv, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a) You can read the data into R with the following script:

&nbsp;`tute1 <- read.csv("tute1.csv", header=TRUE)`

&nbsp;`View(tute1)`

```{r warning=FALSE}
tute1 <- read.csv("tute1.csv", header=TRUE)
View(tute1)
```

b) Convert the data to time series

&nbsp;`mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)`

(The [,-1] removes the first column which contains the quarters as we don’t need them now.)

```{r warning=FALSE}
mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```

c) Construct time series plots of each of the three series

&nbsp;`autoplot(mytimeseries, facets=TRUE)`

Check what happens when you don’t include facets=TRUE.

```{r warning=FALSE}
autoplot(mytimeseries, facets=TRUE)
autoplot(mytimeseries)
```

<hr><hr>

3. Download some monthly Australian retail data from OTexts.org/fpp2/extrafiles/retail.xlsx. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

a) You can read the data into R with the following script:

&nbsp;`retaildata <- readxl::read_excel("retail.xlsx", skip=1)`

The second argument (skip=1) is required because the Excel sheet has two header rows.

```{r warning=FALSE}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

b) Select one of the time series as follows (but replace the column name with your own chosen column):

&nbsp;`myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))`
  
```{r warning=FALSE}
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
```

c) Explore your chosen retail time series using the following functions:

&nbsp;`autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()`

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

```{r warning=FALSE}
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts, lags = 12)
ggAcf(myts)
cat("can see seasonality and trend of the data")
```

<hr><hr>
6. Use the the following graphics functions: `autoplot, ggseasonplot, ggsubseriesplot, gglagplot, ggAcf` and explore features from the following time series: `hsales, usdeaths, bricksq, sunspotarea, gasoline`.
* Can you spot any seasonality, cyclicity and trend?
* What do you learn about the series?

<hr> 
* hsales
```{r warning=FALSE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales, lag.max = 400)
cat("can spot seasonality and cyclicity. The cycle period is about 4 years(100 months)")
```

<hr> 
* usdeaths
```{r warning=FALSE}
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths, lag.max = 60)
cat("can spot seasonality")
```

<hr> 
* bricksq
```{r warning=FALSE}
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq, lag.max = 200)
cat("can spot little seasonality and strong trend")
```

<hr> 
* sunspotarea
```{r warning=FALSE}
autoplot(sunspotarea)
# ggseasonplot(sunspotarea) 
# ggsubseriesplot(sunspotarea)
print("For ggseasonplot, not seasonal | can't draw it. For ggsubseriesplot, not seasonal | useless to draw it")
gglagplot(sunspotarea)
ggAcf(sunspotarea, lag.max = 50)
cat("can spot strong cyclicity")
```

<hr> 
* gasoline
```{r warning=FALSE}
autoplot(gasoline)
ggseasonplot(gasoline)
# ggsubseriesplot(gasoline)
print("The number of weeks is 52 and it looked like it is too much for subseriesplot")
gglagplot(gasoline)
ggAcf(gasoline, lag.max = 1000)
cat("can spot seasonality and trend")
```


